---
title: "Location-regularized BOIN selection: full reproducible analysis (R)"
author: "Siyi Chen"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE,
                      fig.width = 7, fig.height = 4)
source("boin_core.R")     # BOIN conduct + location-regularized selection
source("designs.R")       # 3+3, CRM, biased-coin, k-in-a-row
source("comparators.R")   # logistic-regression / Bayesian model-based selection
```

This single document re-runs every analysis in base R (no external packages, nothing pre-saved)
and regenerates all tables and figures. Set `QUICK <- FALSE` and raise the replicate counts to
match the manuscript exactly; the fast defaults reproduce the same qualitative results in a
handful of minutes. All randomness is seeded.

```{r params}
PHI <- 0.25; N <- 30; J <- 6
QUICK   <- TRUE
n_fixed <- if (QUICK) 4000 else 10000   # reps per fixed scenario   (paper: 10000)
n_rand  <- if (QUICK) 1500 else 10000   # reps per random scenario  (paper: 10000)
n_train <- if (QUICK)   50 else 150     # tuning scenarios
n_test  <- if (QUICK)   80 else 200     # evaluation scenarios
k_pos   <- if (QUICK)   30 else 55      # scenarios per MTD position
k_sweep <- if (QUICK)   40 else 70      # scenarios per sweep cell
n_cross <- if (QUICK)   40 else 110     # scenarios for cross-design / comparators
tab <- elim_table(PHI, N)               # elimination lookup (speed)
SKEL <- c(0.05, 0.12, 0.25, 0.40, 0.55, 0.70)   # CRM skeleton (prior MTD ~ d3)
```

## Cache helpers

```{r helpers}
# Simulate a scenario once under a chosen design; cache the (n,x) tables.
cache_design <- function(p, design, n_sim, seed) {
  set.seed(seed); bnd <- boin_boundaries(PHI); Jl <- length(p)
  nm <- matrix(0L, n_sim, Jl); xm <- matrix(0L, n_sim, Jl); crv <- NULL
  if (design == "crm") crv <- matrix(0, n_sim, Jl)
  own <- integer(n_sim)
  for (t in seq_len(n_sim)) {
    tr <- switch(design,
      boin = simulate_one_trial(p, PHI, N, bnd = bnd, tab = tab),
      "3p3" = sim_3p3(p), bcd = sim_bcd(p, PHI, N),
      krow = sim_krow(p, PHI, N), crm = sim_crm(p, PHI, N, SKEL))
    nm[t, ] <- tr$n; xm[t, ] <- tr$x
    if (design == "crm") crv[t, ] <- tr$curve
    if (design == "3p3") own[t] <- tr$mtd
  }
  list(n = nm, x = xm, curve = crv, own = own, mtd = which.min(abs(p - PHI)))
}
pc_hit <- function(sel_vec, mtd) mean(sel_vec == mtd) * 100
# selection over a cache, returning P_C
pc_std <- function(c) {
  s <- if (is.null(c$curve))
    vapply(seq_len(nrow(c$n)), function(t) select_boin(c$n[t,], c$x[t,], PHI, tab = tab), 0L)
  else vapply(seq_len(nrow(c$n)), function(t) select_curve(c$curve[t,], c$n[t,], PHI), 0L)
  pc_hit(s, c$mtd)
}
pc_prior <- function(c, rho, alpha) {
  s <- if (is.null(c$curve))
    vapply(seq_len(nrow(c$n)), function(t) select_lr(c$n[t,], c$x[t,], PHI, rho, alpha, tab = tab), 0L)
  else vapply(seq_len(nrow(c$n)), function(t) select_curve(c$curve[t,], c$n[t,], PHI, rho, alpha), 0L)
  pc_hit(s, c$mtd)
}
```

## Validation

```{r validation}
tab60 <- elim_table(0.30, 60)
cn <- run_paired(c(0.05,0.15,0.30,0.45,0.60), 0.30, 60, 4000, 0, 2.5, seed = 6, tab = tab60)
cat("rho=0 recovers BOIN exactly:",
    isTRUE(all.equal(cn["boin"], cn["lr"], check.names = FALSE)), "\n")
```

## Table 2 — fixed case-study scenarios, (rho,alpha)=(3.0,2.5)

```{r table2}
FIXED <- list(
 "1 Moderate ramp"=c(0.05,0.10,0.25,0.40,0.55,0.70), "2 Steep cliff"=c(0.02,0.08,0.25,0.55,0.80,0.95),
 "3 Gradual rise"=c(0.08,0.14,0.20,0.26,0.34,0.45),   "5 Sparse DLT"=c(0.04,0.06,0.12,0.25,0.40,0.58),
 "6 Flat near tgt"=c(0.15,0.20,0.24,0.28,0.33,0.40),  "7 Very steep"=c(0.01,0.03,0.25,0.60,0.85,0.95),
 "4 Low-mid"=c(0.12,0.25,0.40,0.55,0.70,0.85), "S1 High MTD"=c(0.02,0.05,0.08,0.15,0.25,0.45),
 "S2 All overdose"=c(0.30,0.45,0.55,0.70,0.80,0.90), "S3 Non-monotone"=c(0.10,0.25,0.15,0.30,0.45,0.60))
t2 <- data.frame()
for (nm in names(FIXED)) {
  r <- run_paired(FIXED[[nm]], PHI, N, n_fixed, 3.0, 2.5, seed = 42, tab = tab)
  t2 <- rbind(t2, data.frame(Scenario=nm, BOIN=round(r["boin"],1),
                             prior=round(r["lr"],1), dPC=round(r["lr"]-r["boin"],1)))
}
rownames(t2) <- NULL; t2
cat("Mean recommended-domain (rows 1-6):", round(mean(t2$dPC[1:6]),1), "pp\n")
```

## Table 3 — out-of-sample tuning and validation

```{r table3}
pool <- gen_random_scenarios(2000, J, PHI, seed = 2024)
interior <- Filter(function(s) s$mtd %in% c(3,4), pool)
set.seed(7); interior <- interior[sample(length(interior))]
train <- interior[seq_len(n_train)]; test <- interior[n_train + seq_len(n_test)]

trc <- lapply(seq_along(train), function(i) cache_design(train[[i]]$p, "boin", n_rand, 5000+i))
grid <- expand.grid(rho=c(2,3,4), alpha=c(2.0,2.5,3.0))
grid$dPC <- apply(grid, 1, function(g)
  mean(sapply(trc, function(c) pc_prior(c, g[1], g[2]) - pc_std(c))))
best <- grid[which.max(grid$dPC), ]

tec <- lapply(seq_along(test), function(i) cache_design(test[[i]]$p, "boin", n_rand, 8000+i))
oos <- function(rho, alpha) {
  d <- sapply(tec, function(c) pc_prior(c, rho, alpha) - pc_std(c))
  c(dPC = mean(d), SE = sd(d)/sqrt(length(d)))
}
t3 <- rbind(
  `tuned`        = c(oos(best$rho, best$alpha), insample = best$dPC),
  `recommended (3.0,2.5)` = c(oos(3.0, 2.5), insample = NA),
  `conservative (2.0,2.0)`= c(oos(2.0, 2.0), insample = NA))
cat("tuned (rho,alpha) =", best$rho, best$alpha, "\n"); round(t3, 2)
```

## Table 4 — parameter sensitivity by true-MTD position

```{r table4}
posp <- gen_random_scenarios(5000, J, PHI, seed = 555)
bypos <- lapply(1:J, function(j) Filter(function(s) s$mtd==j, posp))
cache_pos <- lapply(1:J, function(j) {
  sc <- bypos[[j]]; if (!length(sc)) return(list())
  sc <- sc[seq_len(min(k_pos, length(sc)))]
  lapply(seq_along(sc), function(i) cache_design(sc[[i]]$p, "boin", n_rand, 20000+100*j+i))
})
grp_dpc <- function(js, rho, alpha) {
  cs <- do.call(c, cache_pos[js])
  mean(sapply(cs, function(c) pc_prior(c, rho, alpha) - pc_std(c)))
}
settings <- list(c(1,2.0), c(2,2.0), c(3,2.5), c(3,3.0), c(4,3.0))
t4 <- do.call(rbind, lapply(settings, function(s) data.frame(
  rho=s[1], alpha=s[2],
  interior = round(grp_dpc(c(3,4), s[1], s[2]),1),
  boundary_adj = round(grp_dpc(c(2,5), s[1], s[2]),1),
  boundary = round(grp_dpc(c(1,6), s[1], s[2]),1))))
t4
```

## Table 5 — dose-count, target-rate, and sample-size sweeps

```{r table5}
interior_scen <- function(Jl, phi, seed) {
  lo <- 2; hi <- Jl - 3
  pl <- gen_random_scenarios(3000, Jl, phi, seed)
  sc <- Filter(function(s) s$mtd >= (lo+1) & s$mtd <= (hi+1), pl)  # 1-based interior
  sc[seq_len(min(k_sweep, length(sc)))]
}
sweep_dpc <- function(scen, phi, N, Jl, seed0) {
  tb <- elim_table(phi, N)
  d <- sapply(seq_along(scen), function(i) {
    set.seed(seed0+i); bnd <- boin_boundaries(phi); mtd <- scen[[i]]$mtd; p <- scen[[i]]$p
    hs<-0;hl<-0
    for (t in 1:n_rand) { tr<-simulate_one_trial(p,phi,N,bnd=bnd,tab=tb)
      iso<-iso_estimate(tr$n,tr$x,phi)
      hs<-hs+(select_boin(tr$n,tr$x,phi,iso,tab=tb)==mtd)
      hl<-hl+(select_lr(tr$n,tr$x,phi,3,2.5,iso,tab=tb)==mtd) }
    100*(hl-hs)/n_rand })
  mean(d)
}
Js <- c(5,6,7,8); dJ <- sapply(Js, function(Jl) sweep_dpc(interior_scen(Jl,0.25,1000+Jl),0.25,30,Jl,1e6+Jl))
phis <- c(0.10,0.15,0.20,0.25,0.30,0.35,0.40)
dphi <- sapply(phis, function(ph) sweep_dpc(interior_scen(6,ph,round(2000+ph*100)),ph,30,6,2e6+round(ph*1000)))
Ns <- c(18,24,30,36,42,48); scN <- interior_scen(6,0.25,3030)
dN <- sapply(Ns, function(nn) sweep_dpc(scN,0.25,nn,6,3e6+nn))
cat("J-sweep dPC:  ", paste0(Js,":",round(dJ,1), collapse="  "), "\n")
cat("phi-sweep dPC:", paste0(phis,":",round(dphi,1), collapse="  "), "\n")
cat("N-sweep dPC:  ", paste0(Ns,":",round(dN,1), collapse="  "), "\n")
```

## Table 6 / Figure 5 — transfer across allocation designs

```{r table6}
cross <- interior[seq_len(n_cross)]
designs <- c("3p3","bcd","krow","boin","crm")
t6 <- data.frame(); dpc_cross <- c()
for (dz in designs) {
  cs <- lapply(seq_along(cross), function(i) cache_design(cross[[i]]$p, dz, if (QUICK) 1000 else 3000, 9000+i))
  pcs <- mean(sapply(cs, pc_std)); pcl <- mean(sapply(cs, function(c) pc_prior(c,3.0,2.5)))
  own <- if (dz=="3p3") round(mean(sapply(cs, function(c) pc_hit(c$own, c$mtd))),1) else NA
  t6 <- rbind(t6, data.frame(design=dz, own_rule=own, std=round(pcs,1),
                             prior=round(pcl,1), dPC=round(pcl-pcs,1)))
  dpc_cross[dz] <- pcl - pcs
}
t6
barplot(dpc_cross[c("3p3","bcd","krow","boin","crm")],
        names.arg=c("3+3","BCD","k-in-a-row","BOIN","CRM"),
        col="seagreen", ylab="dPC from location prior (pp)", main="Figure 5: cross-design transfer")
abline(h=0)
```

## Table 7 — model-based selection comparators

```{r table7}
methods <- c("BOIN","logreg (Liao)","bayes-logit (Wakayama)","bayes-cloglog (Wakayama)",
             "BOIN+prior","bayes-logit+prior")
acc <- matrix(0, length(cross), length(methods)); colnames(acc) <- methods
for (i in seq_along(cross)) {
  p <- cross[[i]]$p; mtd <- cross[[i]]$mtd; set.seed(12000+i)
  h <- numeric(length(methods)); nn <- if (QUICK) 800 else 1500
  for (t in 1:nn) {
    tr <- simulate_one_trial(p, PHI, N, tab = tab)
    h[1] <- h[1] + (select_boin(tr$n,tr$x,PHI,tab=tab)==mtd)
    h[2] <- h[2] + (select_logreg(tr$n,tr$x,PHI)==mtd)
    h[3] <- h[3] + (select_bayes(tr$n,tr$x,PHI,"logit")==mtd)
    h[4] <- h[4] + (select_bayes(tr$n,tr$x,PHI,"cloglog")==mtd)
    h[5] <- h[5] + (select_lr(tr$n,tr$x,PHI,3,2.5,tab=tab)==mtd)
    h[6] <- h[6] + (select_bayes(tr$n,tr$x,PHI,"logit",3,2.5)==mtd)
  }
  acc[i, ] <- 100*h/nn
}
pcm <- colMeans(acc)
t7 <- data.frame(method=methods, mean_PC=round(pcm,1), dPC_vs_BOIN=round(pcm-pcm[1],1))
rownames(t7) <- NULL; t7
```

## Figure 1 — location penalty; Figure 3 — gain by position

```{r fig1_3, fig.width=7, fig.height=3.4}
par(mfrow=c(1,2))
z <- seq(0.001,0.999,length=400)
plot(z,(3/J)*1.5*(log(z)+log(1-z)),type="l",lwd=2,col="firebrick",ylim=c(-6,0.5),
     xlab="dose position z", ylab="location prior", main="Figure 1: penalty shape")
lines(z,(2/J)*(log(z)+log(1-z)),lty=2); lines(z,(4/J)*2*(log(z)+log(1-z)),lty=3,col="navy")
zj<-(1:J-0.5)/J; points(zj,(3/J)*1.5*(log(zj)+log(1-zj)),pch=19,col="firebrick")
dpos <- sapply(1:J, function(j) grp_dpc(j, 3.0, 2.5))
barplot(dpos, names.arg=paste0("d",1:J), col="firebrick", main="Figure 3: gain by MTD position",
        ylab="dPC (pp)"); abline(h=0)
```

## Figure 4 — sweeps

```{r fig4, fig.width=9, fig.height=3}
par(mfrow=c(1,3))
plot(Js,dJ,type="b",pch=19,col="navy",xlab="number of doses J",ylab="dPC (pp)",main="(a) J"); abline(h=0,col="grey")
plot(phis,dphi,type="b",pch=15,col="seagreen",xlab="target rate phi",ylab="dPC (pp)",main="(b) phi"); abline(h=0,col="grey")
plot(Ns,dN,type="b",pch=17,col="firebrick",xlab="sample size N",ylab="dPC (pp)",main="(c) N"); abline(h=0,col="grey")
```

Cross-design and comparator implementations are in `designs.R` and `comparators.R`; the core
BOIN conduct and location-regularized selection are in `boin_core.R`.
